---
title: "Poisson and binomial models in INLA"
output: html_notebook
---

Our linear model last time was not a good fit for the data, because we have either count data (Poisson or negtaive binomial distribution) or we have binomal (proportion positive) data. So now, we fit a mixed effect model using poisson distribution


```{r}
library(tidyverse)
library(INLA)

bu<-read_csv("template_BU.csv")

```

When we include a dependency, we need some kind of grouping structure. We dont have one, so I add a random one. This just demonstrates how to code the model

```{r}
bu$Random<-NA
bu$Random[1:45]<-"A"; bu$Random[46:200]<-"B"; bu$Random[201:295]<-"C"
bu$Random<-as.factor(bu$Random)
ggplot(bu)+geom_boxplot(aes(Random, BU_TOT))
```

In our scenario maybe Random is actually a grouping factor of who did the sampling or which lab anaysed the sample, therefore we may suspect that there is dependency.

We will now apply a poisson model with dependency

First, write the model - for no dependency:

BU_TOT_i ~ Poissin(mu) - no hyperparameter
E(BU_Tot_i) = mu_i
Var(BU_Tot_i) = mu_i

mu_i = e^Cov and it is a loglink. Because is e^Cov can only predict positive numbers

First we will do with no dependency then will include dependency

For model with dependency:
BU_TOT_ij ~ Poissin(mu) - no hyperparameter
E(BU_Tot_ij) = mu_ij
Var(BU_Tot_ij) = mu_ij
mu_ij         = Covariate stuff + a_i

 where: 
a_i ~ N(0, sigmaRandom^2)


We now have a hyperparameter in the dependency model



```{r}
f0 <- BU_TOT ~ Forest + Precip_mm  + TOT_POP #no dependency

#dependency
f1 <- BU_TOT ~ Forest + Precip_mm  + TOT_POP +
              f(Random, model = "iid")

I0 <- inla(f0, 
           control.compute = list(dic = TRUE),
           control.predictor = list(compute = TRUE),
           family = "poisson",
           data = bu)

I1 <- inla(f1, 
           control.compute = list(dic = TRUE),
           control.predictor = list(compute = TRUE),
           family = "poisson",
           data = bu)

summary(I0)
summary(I1)


```

For other types of dependency, see the INLA latent models page


